Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers

### 2023-08-18 2_SSL_NE/article/article_main.tex
final edits

Historical improvements in consumer experience metrics for phosphor-converted warm white GaN-based LEDs are shown in \cref{fgr:consumer_experience}. In general illumination applications, a high colour rendering index (CRI) in combination with a specific, tunable range of possible colour temperatures is desirable. Both metrics are determined by LED device emission spectra, which, in turn, depend on the properties of materials used for the conversion of blue light generated by conventional GaN-based LEDs into the white light required for general illumination . It allows us to establish the links between all major improvements in the two consumer experience metrics considered in this study and individual LED innovations  associated either with phosphors or quantum dots used for light down conversion in LEDs. The first commercial white LED produced by Nichia in 1996 used a YAG phosphor that could generate only cool white light \cite{bando1998development}. As shown in \cref{fgr:consumer_experience}, after a series of innovations listed in \cref{tab:innovations}, LEDs today can be tuned for high CRI performance and a range of desirable colour temperatures.

Notably, from detailed descriptions of the history of innovations in this list, provided in Supplementary Note 4, we find that only a single innovation related to LED consumer experience improvements was initially intended exclusively for application in solid-state lighting: the 2016 SALON phosphor compound \cite{Hoerder2019,seibald2019phosphor}. All other innovations in the list were either originally developed for non-LED applications or prominently used knowledge from areas of science and technology beyond LED or SSL (see also \cref{tab:phosphors}). This means that technology spillovers contributed to nearly all consumer experience improvements in LED-based lighting technology, thus playing an important role in widespread adoption of SSL in general illumination applications.

\begin{table}[H]
    \footnotesize
    \centering
    \caption{\textbf{LED innovations related to improvements in consumer experience metrics.}}
    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{Ref.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{NiceTabularX}
    \caption*{Note: \textit{Year} column represents the earliest reported application of innovation in white LEDs. These differ from the years shown in \cref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for LED products that relied on those innovations. Detailed descriptions of the history of each innovation and relevant spectral data are provided in Supplementary Note 4. Desig. = Designation, Ref. - Reference.}
    \label{tab:phosphors}
\end{table}

@@

The structure of our bottom-up white LED manufacturing cost model with process-step resolution is generally based on the 2012 \textit{LEDCOM} cost model \cite{ledcomv2}, but we expand it significantly both in scope and in its ability to capture historical trends. The model captures three historical time periods corresponding to different “eras” in LED manufacturing: the early period of the first high-power white LEDs around 2003, the period of accelerating consumer adoption of LED lighting around 2012, and the most recent period around 2020, the year of our main data collection efforts. For each of these three years, the most prevalent manufacturing equipment was identified through industry periodicals, archived website data from the Internet Archive , and expert interviews. Because the architecture of LED chips has changed significantly since the introduction of the first commercial white LED devices in 1996, three different chip architectures are considered in the model: classical chips, flip chips, and chip-scale package flip chips (see Figure 3 and Supplementary Figures 2-6 for the details of each architecture). The details of the manufacturing process for each architecture were collected from the scientific literature, textbooks  and relevant patents. In addition, two LED life cycle analyses (Scholand and Dillon 2012a)(Casamayor, Su, and Ren 2018) were used to validate the model structure and extract some of the necessary quantitative model inputs. These studies captured a large number of LED manufacturing process steps and included the details on the use of metals, chemicals and electricity for each manufacturing step.

>>>>

Historical improvements in consumer experience metrics for phosphor-converted warm white GaN-based LEDs are shown in \cref{fgr:consumer_experience}. In general illumination applications, a high colour rendering index (CRI) in combination with a specific, tunable range of possible colour temperatures is desirable. Both metrics are determined by LED device emission spectra, which, in turn, depend on the properties of materials used for the conversion of blue light generated by conventional GaN-based LEDs into the white light required for general illumination . It allows us to establish the links between all major improvements in the two consumer experience metrics considered in this study and individual LED innovations  associated either with phosphors or quantum dots used for light down conversion in LEDs. The first commercial white LED produced by Nichia in 1996 used a YAG phosphor that could generate only cool white light \cite{bando1998development}. As shown in \cref{fgr:consumer_experience}, after a series of innovations listed in \cref{tab:phosphors}, LEDs today can be tuned for high CRI performance and a range of desirable colour temperatures.

Notably, from detailed descriptions of the history of innovations in this list, provided in Supplementary Note 4, we find that only a single innovation related to LED consumer experience improvements was initially intended exclusively for application in solid-state lighting: the 2016 SALON phosphor compound \cite{Hoerder2019,seibald2019phosphor}. All other innovations in the list were either originally developed for non-LED applications or prominently used knowledge from areas of science and technology beyond LED or SSL (see also \cref{tab:spillovers}). This means that technology spillovers contributed to nearly all consumer experience improvements in LED-based lighting technology, thus playing an important role in widespread adoption of SSL in general illumination applications.

\begin{table}[H]
    \footnotesize
    \centering
    \caption{\textbf{LED innovations related to improvements in consumer experience metrics.}}
    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{Ref.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{NiceTabularX}
    \caption*{Note: \textit{Year} column represents the earliest reported application of innovation in white LEDs. These differ from the years shown in \cref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for LED products that relied on those innovations. Detailed descriptions of the history of each innovation and relevant spectral data are provided in Supplementary Note 4. Desig.=Designation, Ref.=Reference.}
    \label{tab:phosphors}
\end{table}

@@

The structure of our bottom-up white LED manufacturing cost model with process-step resolution is generally based on the 2012 \textit{LEDCOM} cost model \cite{ledcomv2}, but we expand it significantly both in scope and in its ability to capture historical trends. The model captures three historical time periods corresponding to different “eras” in LED manufacturing: the early period of the first high-power white LEDs around 2003, the period of accelerating consumer adoption of LED lighting around 2012, and the most recent period around 2020, the year of our main data collection efforts. For each of these three years, the most prevalent manufacturing equipment was identified through industry periodicals, archived website data from the Internet Archive , and expert interviews. Because the architecture of LED chips has changed significantly since the introduction of the first commercial white LED devices in 1996, three different chip architectures are considered in the model: classical chips, flip chips, and chip-scale package flip chips (see Figure 3 and Supplementary Figures 2-6 for the details of each architecture). The details of the manufacturing process for each architecture were collected from the scientific literature, textbooks  and relevant patents. In addition, two LED life cycle analyses \cite{Scholand2012}\cite{casamayor2018comparative} were used to validate the model structure and extract some of the necessary quantitative model inputs. These studies captured a large number of LED manufacturing process steps and included the details on the use of metals, chemicals and electricity for each manufacturing step.
